---
title: "Structure and Morphology Population Level MEMA Data"
date: "April 22, 2015"
output: pdf_document
---
##Summary
This is the first experiment to use the proposed LINCs Structure and Morphology staining set in 8 well MEMAs.  Two plates were printed with  35 row by 20 column MEMAs. A 4x7 pin head printed blocks that are 5x5 spots each. Each spot contains one ECM protein paired with Collagen I. The 16 different ECM proteins in the array are in random locations and have an average of 44 replicates.  Each well has the same printed array as shown in the table below.


Barcode|Cells| 1| 2| 3| 4
-----|-----|-----|-----|-----|-----:
LI8V00205|PC3|PBS|HighSerum|EGF|IGF|
LI8V00205|PC3|PBS|EGF|IGF|bFGF
LI8V00206|PC3|PBS|IGF|bFGF|HighSerum
LI8V00206|PC3|PBS|bFGF|HighSerum|EGF


Images of each well were gathered on a Tecan LS Reloaded laser scanner. This staining set includes, DAPI, phalloidin (488nm), CellMask (532 and 555nm) and MitoTracker (635 and 647nm). Data from DAPI staining is only gathered by the Scan^R. 

Tecan data is gathered at the spot population level by fitting round regions of interest (ROIs) to each spot. The Tecan data in this report uses the net values defined as the raw ROI value minus the mean of the local background.


```{r setup, echo=FALSE, message=FALSE}
library("limma")
library("MEMA")
library("data.table")
library("devtools")

```

```{r Read and clean spotmetadata, echo=FALSE, message=FALSE, warnings=FALSE}

#Read in the spot metadata from the gal file
smd <- MEMA::readSpotMetadata("~/Documents/MEMATestExp/LI8V002/Raw Data and Metadata/20150403_LI8V002_16ECM_28pin.gal")
#Make a short name from the content names for labeling in plots
smd$ShortName <- gsub("_.*","",smd$Name)
smd$ShortName <- gsub("-","blank",smd$ShortName)
#Save the spot metadata to a file for use inside the MEMA package
use_data(smd)

#Add the print order and deposition number to the metadata
ldf <- MEMA::readLogData("~/Documents/MEMATestExp/LI8V002/Raw Data and Metadata/20150403-100218.xml")
 #Save the print metadata to a file for use inside the MEMA package
use_data(ldf)
  
spotMetadata <- merge(smd,ldf)
setkey(spotMetadata,Spot)
#Save the merged metadata to a file for use inside the MEMA package
use_data(spotMetadata)  

#Make a rotated version of the spot metadata
spotMetadata180 <- MEMA::rotateMetadata(spotMetadata)
#Save the merged metadata to a file for use inside the MEMA package
use_data(spotMetadata180) 

ARowMetadata <- data.table(spotMetadata,Well=rep(c("A01", "A02","A03","A04"),each=nrow(spotMetadata)))
BRowMetadata <- data.table(spotMetadata180,Well=rep(c("B01", "B02","B03","B04"),each=nrow(spotMetadata180)))

```


```{r Teacan merge, normalize and QA, echo=FALSE, message=FALSE, warnings=FALSE}
#The next steps are to bring in the well metadata, the print order and the Tecan intensity  
#data. The Tecan data includes the raw, background and net data values.
popDataFiles <- c("LI8V00205.txt", "LI8V00206.txt")

barcodes <- gsub(".txt","",popDataFiles)

annDTList <- lapply(barcodes, function(barcode){
  spotsPerWell <- max(spotMetadata$Block) * max(spotMetadata$Row) * max(spotMetadata$Column)
  #Read data from one plate
  popDataRaw<-fread(paste0("~/Documents/MEMATestExp/LI8V002/Raw Data and Metadata/",barcode,".txt"),stringsAsFactors=FALSE)
  #Change the name of the first column
  setnames(popDataRaw,1,"Index")
  #Delete the statistical summary after the population data
  popDataRaw <- popDataRaw[1:spotsPerWell,]
  #Save for use inside the MEMA package
  use_data(popDataRaw) 
  #Organize the data by well
  popData <- MEMA::melt8Well(popDataRaw)
  #Save for use inside the MEMA package
  use_data(popData)
  
  setnames(popData,"Grid","Block")
  popData$Well <- gsub("A","T",popData$Well)
  popData$Well <- gsub("B","A",popData$Well)
  popData$Well <- gsub("T","B",popData$Well)
  setkey(popData, Well, Block, Row, Column)
  
  #Merge in the well metadata
  #Read the well metadata from a multi-sheet Excel file
  wellMetadata <- data.table(MEMA::readMetadata(path.expand(paste0("~/Documents/MEMATestExp/LI8V002/Raw Data and Metadata/",barcode,".xlsx"))), key="Well")  
  popData <- merge(popData, wellMetadata, by="Well")
  #Merge the data with spot metadata that is well row specific
  DT <- merge(popData,rbind(ARowMetadata,BRowMetadata), by = c("Well","Block","Row","Column"), all=TRUE)

  write.table(DT, paste0(barcode,"PopAnn.txt"), sep = "\t",row.names = FALSE, quote=FALSE)
  return(DT)
})
annDT <- rbindlist((annDTList))
#Merge the data with spot metadata that is well row specific
sDT <- merge(annDT,rbind(ARowMetadata,BRowMetadata), by = c("Well","Block","Row","Column"), all=TRUE)

#Randomly assign values of the Net.488 channel to create a null hypothesis result
set.seed(1234)
sDT <- sDT[,Net.488Random := randomizePositions(Net.488),by="Barcode,Well"]

write.table(sDT, paste0("LI8V00205_06PopAnn.txt"), sep = "\t",row.names = FALSE, quote=FALSE)

```
